---
title: "HAPPY HALLOWEEN Candy Miniproject from William Spookyface Bray"
format: gfm
---

```{r}
candy_file <- "candy-data.csv"
candy = read.csv(candy_file, row.names=1)
head(candy)
```
```{r}
nrow(candy)
```
#Question 1) There are 85 different kinds of candy!
```{r}
table(candy$fruity)
```
# Question 2) There are 38 different kinds of fruity candy (in this case, the '1' is tallying the presence of candy!
```{r}
candy["Twix", ]$winpercent
```
#cool example
```{r}
candy["Charleston Chew", ]$winpercent
```
#Question 3: I'm claiming/lying that Charleston Chew is my favorite candy, even though it's only really palatable when frozen. It has a win percentage of 38.975 
```{r}
candy["Kit Kat", ]$winpercent
```
#Question 4: 76.77% win or so..
```{r}
candy["Tootsie Roll Snack Bars", ]$winpercent
```
#Question 5: 49.635
```{r}
library(dplyr)
candy |> 
  filter(rownames(candy) %in% c("Dum Dums", "Twix")) |>
select(winpercent)
```
```{r}
library("skimr")
skim(candy)
```
```{r}
c("j", "K", "l") %in% c("x", "K", "d")
```
#cool function bro!
```{r}
candy |> 
  filter(winpercent > 75) |>
  filter(pricepercent < 0.5)
```
#nice!

#Question 6: The win percentage is quite large relative to the other columns

#Question 7: The '0' and '1' are the absence or presence of chocolate, respectively!
```{r}
hist(candy$winpercent)
```
#Question 8: two different ways; ggplot is the best
```{r}
library(ggplot2)
ggplot(candy, aes(winpercent)) + geom_histogram(bins = 10) + theme_dark()
```
#Question 9: the distribution doesn't appear to be symmetrical
```{r}
summary(candy$winpercent)
```
#Question 10 The center of the distribution is in fact below 50
```{r}
candy$winpercent[as.logical(candy$chocolate)]
```
```{r}
candy$winpercent[as.logical(candy$fruity)]
```
```{r}
inds <- as.logical(candy$chocolate)
candy[inds,]$winpecent
```
#huh
```{r}
summary(candy$winpercent[as.logical(candy$chocolate)])
```

```{r}
summary(candy$winpercent[as.logical(candy$fruity)])
```
#Question 11: chocolate is definitely preferred!
```{r}
t.test(summary(candy$winpercent[as.logical(candy$fruity)]), summary(candy$winpercent[as.logical(candy$chocolate)]))
```
#hmmm, wrong 
```{r}
inds <- candy$chocolate ==1
choc.win <- candy[inds,]$winpercent

inds <- candy$fruity ==1
fruit.win <- candy[inds,]$winpercent
```

```{r}
summary(fruit.win)
summary(choc.win)
```
```{r}
t.test(fruit.win, choc.win)
```
#There we go! 
#Question12: yes indeed it is significant
```{r}
play <- c(2,1,5,3)
sort(play)
```
```{r}
play[order(play)]
```
```{r}
n <- c("d", "a")
n [order(n)]
```
#Cool function to remember!
```{r}
sort(candy$winpercent)
```
#I see, just giving me the win percent!
```{r}
inds <- order(candy$winpercent)
head(candy[inds,])
```
#Okay so that's working fine...
#Question 13: The five least liked candies are Nik L Nip, Boston Baked Beans, Chiclet, Super Bubble, Jawbusters, Root Beer Barrels
```{r}
inds <- order(candy$winpercent)
head(candy[rev(inds),])
```
# Question 14 Reese's peanut butter and miniatures take the top two, followed by Twix, Kit Kat, Snickers and Reese's Pieces
#question 15:
```{r}
library(ggplot2)

ggplot(candy) + 
  aes(winpercent, rownames(candy)) +
  geom_col()
```
#Doesn't look too great
#Question 16:
```{r}
ggplot(candy) + aes(x=winpercent, y=reorder(rownames(candy), winpercent)) + geom_col()
```
#Hey that's in order now
# color your favorite candy!
```{r}
my_cols=rep("black", nrow(candy))
my_cols[as.logical(candy$chocolate)] = "brown"
my_cols[as.logical(candy$bar)] = "purple"
my_cols[as.logical(candy$fruity)] = "pink"
my_cols[rownames(candy)=="Charleston Chew"] = "orange"

ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  geom_col(fill=my_cols) 
```
# If an individual (Charleston Chew) assignment came first it woud be overwritten, so the final bit should be last!

#Question 17: Sixlets are the least popular form of chocolate, which is not unfair
#Question 18: Starburst are justifiably beloved as the most popular of otherwise pretty lousy fruity candies

#Geom_text_repel is super useful below to separate
```{r}
library(ggrepel)
ggplot(candy) +
  aes(winpercent, pricepercent, label=rownames(candy)) +
  geom_point(col=my_cols) + 
  geom_text_repel(col=my_cols, size=3.3, max.overlaps = 12)
```
#still a bit overplotted, but improved!

#Question 19: Reese's miniatures are a solid bet here

#Question 20: Nik L Nip remains the least popula,yet most expensive

#ooh a unique sorting situation here...
```{r}
ord <- order(candy$pricepercent, decreasing = TRUE)
head( candy[ord,c(11,12)], n=5 )
```
# try corrplot to examine correlation structure
#
```{r}
ggplot(candy) +
  aes(pricepercent, reorder(rownames(candy), pricepercent)) +
  geom_segment(aes(yend = reorder(rownames(candy), pricepercent), 
                   xend = 0), col="blue") +
    geom_point()
```


```{r}
library(corrplot)
cij <- cor(candy)

corrplot(cij)

```
#how do we read this? -1 anticorrelated, +1 is positive correlation; so fruity is anticorrelated with chocolate
#make a pca plot of the situation

#Question 22: Fruity and chocolate are quiet thoroughly anticorrelated
#Question 23: Chocolate and win percentage are quite highly positively correlated; per the hint chocolate fruity candies are fairly rare, and so not popular

```{r}
pca <- prcomp(candy, scale = TRUE)
summary(pca)

```

```{r}
plot(pca$x[,1], pca$x[,2], col=my_cols)
```
```{r}
pca$rotation[,1]
```
```{r}
my_data <- cbind(candy, pca$x[,1:3])
```
```{r}
w <- ggplot(my_data) + 
        aes(x=PC1, y=PC2, 
            size=winpercent/100,  
            text=rownames(my_data),
            label=rownames(my_data)) +
        geom_point(col=my_cols)

w
```
```{r}
library(ggrepel)

w + geom_text_repel(size=3.3, col=my_cols, max.overlaps = 7)  + 
  theme(legend.position = "none") +
  labs(title="Halloween Candy PCA",
       subtitle="Colored by type: chocolate bar (dark brown), chocolate other (light brown), fruity (red), other (black)",
       caption="Data from 538")
```
```{r}
par(mar=c(8,4,2,2))
barplot(pca$rotation[,1], las=2, ylab="PC1 Contribution")
```
#Question 24: Fruity, hard, and pluribus; this makes perfect sense, given chocolate goes precisely the opposite direction of fruity, and fruity candies are often sold in bag form...
#Plotly is pretty cool, but alas, have to comment out

#```{r}
#library(plotly)
#ggplotly(w)
```

